---
title: "EIS 2026 Information Sources"
subtitle: "Where voters get election news, where they look for specific kinds of election information, and how both compare to 2024"
date: "`r paste0('Last updated ', format(Sys.time(), '%B '), as.integer(format(Sys.time(), '%d')), format(Sys.time(), ', %Y at '), sub('^0', '', format(Sys.time(), '%I:%M %p')))`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: readable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, dpi = 96)
knitr::opts_knit$set(upload.fun = knitr::image_uri)   # embed plots (self-contained, no pandoc)
suppressMessages({library(tidyverse); library(survey); library(knitr)})
theme_set(theme_minimal(base_size = 12))

# Colors. BAR_COLOR/RANK_COLOR/PID_COLORS are for Parts 1-2 (2026 only); YEAR_COLORS is for Part 3
# (the 2024-2026 trend) and is copied verbatim from eis_2026_05_trend_comparison_report.Rmd so a
# reader sees the same year-to-color mapping in both reports.
BAR_COLOR   <- "#2a78d6"
RANK_COLOR  <- "#eb6834"
PID_COLORS  <- c(Dem = "#2a78d6", Ind = "#898781", Rep = "#d03b3b")
YEAR_COLORS <- c("2024" = "#2a78d6", "2026" = "#eb6834")
```

```{r load-data}
data         <- readRDS("output/eis_2026_clean.rds")
var.index    <- read.csv("output/eis_2026_variable_index.csv", stringsAsFactors = FALSE)
sources.long <- readRDS("output/eis_2026_sources_long.rds") %>%
  left_join(data %>% select(resp_id, pid3), by = "resp_id") %>%
  mutate(display = str_trim(str_extract(item_label, "(?<=: ).*")))
cum          <- readRDS("output/eis_cumulative.rds") %>%
  mutate(year = factor(year, levels = c(2024, 2026)))
```

```{r part3-setup}
# Every {need}_src_* / src_* column built in eis_2026_04_cumulative_data.R's E.2b arrives as a
# 2-level factor ("0"/"1"), the same convention the existing trend report's own helpers-2 chunk
# already converts before use.
new_src_cols <- c(
  paste0("src_", c("tv_local","tv_national","radio","news_local","news_national","social",
                    "search","podcast","newsletter","aggregator","chatbot","friends_family")),
  paste0(rep(c("reg_", "run_", "won_"), each = 6),
         rep(c("src_search","src_chatbot","src_fact_checking","src_elected_officials",
               "src_commentator","src_civic_religious"), 3))
)
cum <- cum %>% mutate(across(all_of(new_src_cols), ~ as.numeric(as.character(.x))))

NEW_LABELS     <- c(search = "Online search engine", chatbot = "AI-enabled chatbot")
DROPPED_LABELS <- c(fact_checking = "Fact-checking organizations", elected_officials = "Elected officials generally",
                    commentator = "Favorite commentator/analyst", civic_religious = "Civic/religious organizations")

# Prints the "new in 2026" / "dropped after 2024" bar charts for one info-need battery (reg/run/won,
# each of which has 2 new and 4 dropped items - both get a chart). Called with results='asis'.
render_new_dropped <- function(need) {
  new_cols <- paste0(need, "_src_", names(NEW_LABELS))
  new_props <- map_dfr(new_cols, function(col) svy_prop(cum %>% filter(year == 2026), col, "weight_common") %>%
                          mutate(item = NEW_LABELS[[str_remove(col, paste0("^", need, "_src_"))]], .before = 1))
  cat("\n\n**New in 2026 (no 2024 counterpart)**\n\n")
  print(plot_battery_bar(new_props, x_label = "Weighted % selected, 2026 only"))

  dropped_cols <- paste0(need, "_src_", names(DROPPED_LABELS))
  dropped_props <- map_dfr(dropped_cols, function(col) svy_prop(cum %>% filter(year == 2024), col, "weight_common") %>%
                              mutate(item = DROPPED_LABELS[[str_remove(col, paste0("^", need, "_src_"))]], .before = 1))
  cat("\n\n**Dropped after 2024 (no 2026 counterpart)**\n\n")
  print(plot_battery_bar(dropped_props, x_label = "Weighted % selected, 2024 only"))
}
```

```{r pool-official-mode}
# Contact mode is a single nominal item, not a multiselect battery, so it is the one variable in
# this report that is not already poolable by filtering eis_2026_sources_long.rds - it never went
# into that long table to begin with. The hypothetical arm's wording is "I would visit/call/visit"
# vs. the actual arm's "I visit/call/visit" (confirmed directly against the data while building this
# report); normalizing that prefix away before coalescing keeps the pooled factor at 5 levels
# instead of silently ballooning to 8.
normalize_mode <- function(x) str_replace(as.character(x), "^I would ", "I ")

data <- data %>%
  mutate(
    reg_official_mode_pooled = factor(coalesce(normalize_mode(reg_act_official_mode), normalize_mode(reg_hyp_official_mode)),
                                       levels = levels(reg_act_official_mode)),
    run_official_mode_pooled = factor(coalesce(normalize_mode(run_act_official_mode), normalize_mode(run_hyp_official_mode)),
                                       levels = levels(run_act_official_mode)),
    won_official_mode_pooled = factor(coalesce(normalize_mode(won_act_official_mode), normalize_mode(won_hyp_official_mode)),
                                       levels = levels(won_act_official_mode))
  )
```

```{r helpers-wide, include=FALSE}
# Estimators and plot builders for a single wide column or a battery of wide columns - ported
# unchanged from eis_2026_06_design_data.R / eis_2026_07_plots_overview.R, used throughout Part 1
# and for the Part 2 drill-downs that have no arm to pool (contact mode).

factor_props <- function(data, factor_col, weight_col = "weight") {
  sub <- data %>% filter(!is.na(.data[[factor_col]]))
  des <- svydesign(ids = ~1, weights = reformulate(weight_col), data = sub)
  est <- svymean(reformulate(factor_col), des)
  tibble(category = sub(paste0("^", factor_col), "", names(est)),
         pct      = as.numeric(est) * 100,
         ci_low   = confint(est)[, 1] * 100,
         ci_high  = confint(est)[, 2] * 100,
         n        = nrow(sub))
}

svy_prop <- function(data, indicator_col, weight_col = "weight") {
  sub <- data %>% filter(!is.na(.data[[indicator_col]]))
  des <- svydesign(ids = ~1, weights = reformulate(weight_col), data = sub)
  est <- svyciprop(reformulate(indicator_col), des, method = "logit")
  tibble(pct = as.numeric(est) * 100, ci_low = confint(est)[1] * 100, ci_high = confint(est)[2] * 100, n = nrow(sub))
}

battery_props <- function(data, cols, weight_col = "weight") {
  data.bin <- data %>% mutate(across(all_of(cols), ~ as.numeric(as.character(.x))))
  map_dfr(cols, function(col) svy_prop(data.bin, col, weight_col) %>% mutate(item = col, .before = 1))
}

factor_props_by <- function(data, factor_col, by_col, weight_col = "weight") {
  data %>%
    filter(!is.na(.data[[by_col]])) %>%
    group_by(.data[[by_col]]) %>%
    group_modify(~ factor_props(.x, factor_col, weight_col)) %>%
    ungroup()
}

item_display_labels <- function(cols) {
  var.index %>% filter(var_name %in% cols) %>%
    transmute(var_name, display = str_trim(str_extract(var_label, "(?<=: ).*"))) %>%
    deframe()
}

rank_display_labels <- function(rank_1st_col) {
  parent.qid <- var.index %>% filter(var_name == rank_1st_col) %>% pull(derived_from) %>% str_extract("BPC[0-9]+")
  var.index %>% filter(base == parent.qid, var_type == "multiselect") %>%
    transmute(item_tag, display = str_trim(str_extract(var_label, "(?<=: ).*"))) %>%
    deframe()
}
```

```{r helpers-long, include=FALSE}
# Estimators built on eis_2026_sources_long.rds instead of wide columns - used throughout Part 2,
# where "which arm(s) count as in scope" changes per chart (exclude never-seekers / include everyone
# / never-seekers only) but is always expressible as a filter on the long table's `battery` column,
# with no new pooled wide column required. "Include everyone" filters to BOTH arms' battery values at
# once; since exactly one arm is `in_base` per respondent, that filter alone already is the pooled
# population - no coalesce() needed.

prep_battery_long <- function(battery_values) {
  sources.long %>% filter(battery %in% battery_values, in_base)
}

# One weighted proportion for one already-filtered subset - the same svyciprop() logit pattern as
# svy_prop() above, reused directly rather than via srvyr/svyby.
prop_one_group <- function(sub, indicator_col, weight_col = "weight") {
  des <- svydesign(ids = ~1, weights = reformulate(weight_col), data = sub)
  est <- svyciprop(reformulate(indicator_col), des, method = "logit")
  tibble(pct = as.numeric(est) * 100, ci_low = confint(est)[1] * 100, ci_high = confint(est)[2] * 100, n = nrow(sub))
}

# `indicator_col` names a logical column the caller has already added to `long_data` (e.g.
# `selected == "1"`, or `coalesce(rank == 1, FALSE)`). Set by_party = FALSE to collapse to one
# weighted % per item overall; TRUE also splits by pid3.
#
# An explicit loop over item_tag (x pid3), calling svyciprop() fresh per subset, rather than
# group_by()/summarise(survey_mean()) or svyby() - the same recipe eis_2026_06_design_data.R's
# svy_prop_by()/factor_props_by() already use for "estimate X within each level of a grouping
# variable" elsewhere in this project. Verified to reproduce the grouped-estimator numbers exactly
# (0 difference) against the established wide-column pattern before use below.
battery_props_by_party <- function(long_data, indicator_col, weight_col = "weight", by_party = TRUE) {
  if (by_party) long_data <- long_data %>% filter(!is.na(pid3))
  tags <- sort(unique(as.character(long_data$item_tag)))
  results <- list()
  k <- 1
  if (by_party) {
    for (tg in tags) for (p in levels(long_data$pid3)) {
      sub <- long_data %>% filter(item_tag == tg, pid3 == p)
      if (nrow(sub) == 0) next
      results[[k]] <- prop_one_group(sub, indicator_col, weight_col) %>%
        mutate(item_tag = tg, item = sub$display[1], pid3 = p, .before = 1)
      k <- k + 1
    }
  } else {
    for (tg in tags) {
      sub <- long_data %>% filter(item_tag == tg)
      if (nrow(sub) == 0) next
      results[[k]] <- prop_one_group(sub, indicator_col, weight_col) %>%
        mutate(item_tag = tg, item = sub$display[1], .before = 1)
      k <- k + 1
    }
  }
  bind_rows(results)
}
```

```{r plots, include=FALSE}
# Bar chart for one factor's weighted distribution (nominal items - e.g. contact mode). Ported from
# eis_2026_07_plots_overview.R.
plot_categorical <- function(props, x_label = "Weighted %", order = NULL) {
  props <- if (is.null(order)) props %>% mutate(category = fct_reorder(category, pct))
           else                props %>% mutate(category = factor(category, levels = order))
  ggplot(props, aes(x = pct, y = category)) +
    geom_col(fill = BAR_COLOR, width = 0.65) +
    geom_linerange(aes(xmin = ci_low, xmax = ci_high), linewidth = 0.9) +
    scale_x_continuous(expand = expansion(mult = c(0, .06))) +
    labs(x = x_label, y = NULL)
}

# Horizontal bar chart for a whole battery, sorted by size. Ported from eis_2026_07_plots_overview.R
# (`item_labels` optional - the long-table helpers above already put display text in `item`).
plot_battery_bar <- function(battery, item_labels = NULL, x_label = "Weighted %") {
  if (!is.null(item_labels)) battery <- battery %>% mutate(item = unname(item_labels[item]))
  battery <- battery %>% mutate(item = fct_reorder(item, pct))
  ggplot(battery, aes(x = pct, y = item)) +
    geom_col(fill = BAR_COLOR, width = 0.65) +
    geom_linerange(aes(xmin = ci_low, xmax = ci_high), linewidth = 0.9) +
    scale_x_continuous(expand = expansion(mult = c(0, .06))) +
    labs(x = x_label, y = NULL)
}

# NEW. A battery's weighted % selected (bar) with an optional ranked-#1 % overlay (point) on the
# same axis and category order - both are proportions of the same base, so they are directly
# comparable. Pass rank_props = NULL to omit the overlay (used for Part 1's drill-down batteries,
# none of which have a rank follow-up).
plot_combo_bar_point <- function(sel_props, rank_props = NULL, x_label = "Weighted %") {
  ord <- sel_props %>% arrange(pct) %>% pull(item)
  sel_props <- sel_props %>% mutate(item = factor(item, levels = ord))
  p <- ggplot(sel_props, aes(y = item)) +
    geom_col(aes(x = pct, fill = "Selected (top 3)"), width = 0.65) +
    geom_linerange(aes(xmin = ci_low, xmax = ci_high), linewidth = 0.9) +
    scale_fill_manual(values = c("Selected (top 3)" = BAR_COLOR), name = NULL)
  if (!is.null(rank_props)) {
    rank_props <- rank_props %>% mutate(item = factor(item, levels = ord))
    p <- p + geom_point(data = rank_props, aes(x = pct, color = "Ranked #1"), size = 2.6, shape = 18) +
      scale_color_manual(values = c("Ranked #1" = RANK_COLOR), name = NULL)
  }
  p + scale_x_continuous(expand = expansion(mult = c(0, .06))) +
    labs(x = x_label, y = NULL) +
    theme(legend.position = "top")
}

# NEW. `props_by` is battery_props_by_party()'s by-party output (item, pid3, pct, ci_low, ci_high).
# Dodged, not stacked - each item's selection is an independent 0/1, not a mutually-exclusive
# category, so the three parties' bars for one item do not sum to anything meaningful.
plot_battery_dodge_party <- function(props_by, x_label = "Weighted %") {
  ord <- props_by %>% group_by(item) %>% summarise(m = mean(pct), .groups = "drop") %>% arrange(m) %>% pull(item)
  props_by <- props_by %>% mutate(item = factor(item, levels = ord))
  ggplot(props_by, aes(x = pct, y = item, fill = pid3)) +
    geom_col(position = position_dodge(width = 0.75), width = 0.7) +
    geom_linerange(aes(xmin = ci_low, xmax = ci_high), position = position_dodge(width = 0.75), linewidth = 0.7) +
    scale_fill_manual(values = PID_COLORS, name = NULL) +
    scale_x_continuous(expand = expansion(mult = c(0, .06))) +
    labs(x = x_label, y = NULL) +
    theme(legend.position = "top")
}

# Single 100%-stacked bar per subgroup - for a genuinely mutually-exclusive nominal item (contact
# mode) broken out by party. Ported from eis_2026_06_design_data.R Part D.6.
plot_prop_stack_by <- function(props_by, by_col, colors, x_label = "Weighted %") {
  props_by <- props_by %>% mutate(category = factor(category, levels = names(colors)))
  ggplot(props_by, aes(x = pct, y = .data[[by_col]], fill = category)) +
    geom_col(width = 0.65) +
    scale_fill_manual(values = colors, name = NULL) +
    scale_x_continuous(expand = expansion(mult = c(0, .02))) +
    labs(x = x_label, y = NULL) +
    theme(legend.position = "top", panel.grid = element_blank())
}
```

```{r helpers-trend, include=FALSE}
# Ported unchanged from eis_2026_05_trend_comparison_report.Rmd, for Part 3.
prop_by_year <- function(data, indicator_col, weight_col = "weight_common") {
  map_dfr(levels(data$year), function(yr) {
    sub <- data %>% filter(year == yr, !is.na(.data[[indicator_col]]))
    des <- svydesign(ids = ~1, weights = reformulate(weight_col), data = sub)
    est <- svyciprop(reformulate(indicator_col), des, method = "logit")
    tibble(year = yr, pct = as.numeric(est) * 100,
           ci_low = confint(est)[1] * 100, ci_high = confint(est)[2] * 100, n = nrow(sub))
  })
}

prop_table_by_item <- function(data, cols, weight_col = "weight_common") {
  map_dfr(cols, function(col) prop_by_year(data, col, weight_col) %>% mutate(item = col, .before = 1))
}

plot_dumbbell <- function(df, x_label, show_ci = TRUE, sort_desc_change = TRUE) {
  wide <- df %>% select(item, year, pct) %>% pivot_wider(names_from = year, values_from = pct, names_prefix = "y")
  ord <- if (sort_desc_change) {
    wide %>% mutate(chg = abs(y2026 - y2024)) %>% arrange(chg) %>% pull(item)
  } else {
    wide %>% arrange(y2026) %>% pull(item)
  }
  df <- df %>% mutate(item = factor(item, levels = ord))

  p <- ggplot(df, aes(x = pct, y = item)) +
    geom_line(aes(group = item), color = "grey75", linewidth = 0.6)
  if (show_ci) {
    p <- p + geom_linerange(aes(xmin = ci_low, xmax = ci_high, color = year), linewidth = 2, alpha = 0.35)
  }
  yr_levels <- names(YEAR_COLORS)
  p <- p +
    geom_point(data = df %>% filter(year == yr_levels[1]), aes(color = year), size = 4.2, shape = 1, stroke = 1.4) +
    geom_point(data = df %>% filter(year == yr_levels[2]), aes(color = year), size = 3, shape = 16)
  p + scale_color_manual(values = YEAR_COLORS, name = NULL) +
    labs(x = x_label, y = NULL) +
    theme(legend.position = "top", panel.grid.minor = element_blank())
}
```

## Methodology

This report covers where 2026 registered voters get election news and information, where they look for three specific kinds of election information (how to register and vote, how elections are run, who won), and how both compare to the 2024 field. Parts 1-2 use `weight`, the delivered 2026 Morning Consult weight; Part 3 uses `weight_common`, which re-rakes both years to one shared demographic target so a year-over-year difference reflects a real difference between respondents rather than a difference in what each year's own weight was correcting toward (see `eis_2026_06_design_data.R` and `eis_2026_05_trend_comparison_report.Rmd` for the fuller argument). Every proportion below carries a 95% confidence interval from `survey::svyciprop()` (logit method); a rank has no standard error of its own and is shown without one.

# Part 1: Where Voters Get Election News

## Main sources

```{r part1-main, fig.height=4.5}
src_cols <- var.index %>% filter(base == "BPC2", var_type == "multiselect") %>% pull(var_name)

sel_props <- data %>%
  mutate(across(all_of(src_cols), ~ as.numeric(as.character(.x)))) %>%
  battery_props(src_cols) %>%
  mutate(item = unname(item_display_labels(src_cols)[item]))

rank_props <- factor_props(data, "src_rank_1st") %>%
  mutate(item = unname(rank_display_labels("src_rank_1st")[category])) %>%
  select(item, pct, ci_low, ci_high, n)

plot_combo_bar_point(sel_props, rank_props, x_label = "Weighted %")
```

n = `r format(sel_props$n[1], big.mark = ",")`. Bars show the weighted % selecting each source among their top three; the diamond shows the weighted % who ranked that source their #1 choice.

## Drill-downs

```{r part1-drilldowns, fig.height=3.5}
render_drilldown <- function(base_qid, section_title) {
  cols <- var.index %>% filter(base == base_qid, var_type == "multiselect") %>% pull(var_name)
  base_cond <- var.index %>% filter(var_name == cols[1]) %>% pull(base_condition)
  props <- data %>% mutate(across(all_of(cols), ~ as.numeric(as.character(.x)))) %>% battery_props(cols)
  cat("\n\n###", section_title, "\n\n")
  print(plot_battery_bar(props, item_labels = item_display_labels(cols), x_label = "Weighted %"))
  cat("\n\nn =", format(props$n[1], big.mark = ","), "— Base:", base_cond, "\n\n")
}
```

```{r part1-tv, results='asis', fig.height=3}
render_drilldown("BPC3", "National television networks")
```

```{r part1-radio, results='asis', fig.height=2.5}
render_drilldown("BPC4", "Radio programming")
```

```{r part1-outlet, results='asis', fig.height=6}
render_drilldown("BPC5", "Print and online news outlets")
```

```{r part1-social, results='asis', fig.height=4}
render_drilldown("BPC6", "Social media platforms")
```

```{r part1-bot, results='asis', fig.height=3}
render_drilldown("BPC10", "AI chatbots")
```

# Part 2: Where Voters Look for Specific Kinds of Information

```{r part2-helpers, include=FALSE}
NEED_MODE_COL <- c(reg = "reg_official_mode_pooled", run = "run_official_mode_pooled", won = "won_official_mode_pooled")
DRILLDOWN_TITLES <- c(social = "Social media platform", bot = "AI chatbot")

render_need_section <- function(need, need_title) {
  arms <- tribble(
    ~arm_title,               ~battery_values,
    "Exclude never-seekers",  paste0(need, "_act"),
    "Include everyone",       c(paste0(need, "_act"), paste0(need, "_hyp")),
    "Never-seekers only",     paste0(need, "_hyp")
  )

  cat("\n\n## ", need_title, "\n\n", sep = "")
  cat("\n\n### Which sources\n\n")

  for (i in seq_len(nrow(arms))) {
    cat("\n\n#### ", arms$arm_title[i], "\n\n", sep = "")
    long <- prep_battery_long(arms$battery_values[[i]])

    sel <- battery_props_by_party(long %>% mutate(is_selected = selected == "1"), "is_selected", by_party = FALSE)
    rnk <- battery_props_by_party(long %>% mutate(is_first = coalesce(rank == 1, FALSE)), "is_first", by_party = FALSE)
    print(plot_combo_bar_point(sel, rnk, x_label = "Weighted %"))
    cat("\n\nn =", format(sel$n[1], big.mark = ","), "\n\n")

    sel_party <- battery_props_by_party(long %>% mutate(is_selected = selected == "1"), "is_selected", by_party = TRUE)
    rnk_party <- battery_props_by_party(long %>% mutate(is_first = coalesce(rank == 1, FALSE)), "is_first", by_party = TRUE)
    cat("\n\n**By party ID — selected**\n\n")
    print(plot_battery_dodge_party(sel_party, x_label = "Weighted %"))
    cat("\n\n**By party ID — ranked #1**\n\n")
    print(plot_battery_dodge_party(rnk_party, x_label = "Weighted %"))
  }

  cat("\n\n### Drill-downs (everyone included)\n\n")
  cat("\n\n#### Contact mode (local/state election officials)\n\n")

  mode_col <- NEED_MODE_COL[[need]]
  mode_overall <- factor_props(data, mode_col)
  print(plot_categorical(mode_overall, x_label = "Weighted %"))
  cat("\n\nn =", format(mode_overall$n[1], big.mark = ","), "\n\n")

  mode_levels <- levels(data[[mode_col]])
  mode_colors <- setNames(colorRampPalette(c("#b7d3f6", "#0d366b"))(length(mode_levels)), mode_levels)
  cat("\n\n**By party ID**\n\n")
  print(plot_prop_stack_by(factor_props_by(data, mode_col, "pid3"), by_col = "pid3", colors = mode_colors))

  for (dd in names(DRILLDOWN_TITLES)) {
    cat("\n\n#### ", DRILLDOWN_TITLES[[dd]], "\n\n", sep = "")
    long_dd <- prep_battery_long(paste0(need, c("_act_", "_hyp_"), dd))

    sel_dd <- battery_props_by_party(long_dd %>% mutate(is_selected = selected == "1"), "is_selected", by_party = FALSE)
    print(plot_battery_bar(sel_dd, x_label = "Weighted %"))
    cat("\n\nn =", format(sel_dd$n[1], big.mark = ","), "\n\n")

    sel_dd_party <- battery_props_by_party(long_dd %>% mutate(is_selected = selected == "1"), "is_selected", by_party = TRUE)
    cat("\n\n**By party ID**\n\n")
    print(plot_battery_dodge_party(sel_dd_party, x_label = "Weighted %"))
  }
}
```

```{r part2-reg, results='asis', fig.height=4.5, fig.width=10}
render_need_section("reg", "Registering and Voting")
```

```{r part2-run, results='asis', fig.height=4.5, fig.width=10}
render_need_section("run", "How Elections Are Run")
```

```{r part2-won, results='asis', fig.height=4.5, fig.width=10}
render_need_section("won", "Who Won an Election")
```

# Part 3: 2026 vs. 2024 Trend

Scoped to the election-information-source questions covered in Parts 1-2. For every other survey topic (right direction, AI attitudes, confidence, concern, etc.), see `eis_2026_05_trend_comparison_report.Rmd`. All estimates below use `weight_common`.

## Main election-news sources (BPC2)

```{r part3-bpc2-common, fig.height=4}
bpc2_items <- c(tv_local = "Local/regional TV", tv_national = "National TV", radio = "Radio",
                news_local = "Local/state print-online", news_national = "National print-online",
                social = "Social media", search = "Search engine", podcast = "Podcasts",
                newsletter = "Newsletters/blogs/forums", aggregator = "News aggregator apps",
                chatbot = "AI chatbot")
bpc2_cols <- paste0("src_", names(bpc2_items))
bpc2_trend <- prop_table_by_item(cum, bpc2_cols) %>%
  mutate(item = bpc2_items[str_remove(item, "^src_")])
plot_dumbbell(bpc2_trend, "Weighted % selected (among top 3)")
```

```{r part3-bpc2-new, results='asis'}
ff <- svy_prop(cum %>% filter(year == 2026), "src_friends_family", "weight_common")
cat(sprintf("\n\n**New in 2026:** %.1f%% (95%% CI %.1f-%.1f) of 2026 respondents selected \"Friends and/or family\" as a main news source — this option did not exist in the 2024 survey. There is no item in the 2024 survey that was dropped from this battery.\n\n",
            ff$pct, ff$ci_low, ff$ci_high))
```

## Registering and voting

```{r part3-reg, fig.height=3}
reg_items <- c(local_officials = "Local/county officials", state_officials = "State officials",
               federal_site = "Federal agency/website", news_media = "News media",
               social = "Social media influencer", friends_family = "Friends/family",
               advocacy = "Election-integrity/advocacy orgs", campaign = "Candidate, campaign, or party")
reg_trend <- prop_table_by_item(cum, paste0("reg_src_", names(reg_items))) %>%
  mutate(item = reg_items[str_remove(item, "^reg_src_")])
plot_dumbbell(reg_trend, "Weighted % selected (among top 3)")
```

```{r part3-reg-nd, results='asis', fig.height=2.5}
render_new_dropped("reg")
```

## How elections are run

```{r part3-run, fig.height=3}
run_trend <- prop_table_by_item(cum, paste0("run_src_", names(reg_items))) %>%
  mutate(item = reg_items[str_remove(item, "^run_src_")])
plot_dumbbell(run_trend, "Weighted % selected (among top 3)")
```

```{r part3-run-nd, results='asis', fig.height=2.5}
render_new_dropped("run")
```

## Who won

```{r part3-won, fig.height=3}
won_trend <- prop_table_by_item(cum, paste0("won_src_", names(reg_items))) %>%
  mutate(item = reg_items[str_remove(item, "^won_src_")])
plot_dumbbell(won_trend, "Weighted % selected (among top 3)")
```

```{r part3-won-nd, results='asis', fig.height=2.5}
render_new_dropped("won")
```
